---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(rmarkdown)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
#devtools::build_readme() 
```

# sExtinct

This package provides functions to infer the extinction status of species based on the distribution of historic sightings (and/or abundances). This combines several sighting based estimators of historical extinction, allowing them to be run simultaneously or individually. Code for this package was contributed by Ben Collen, Gene Hunt and Tracy Rout. Additional code was taken from McPherson & Myers (2009).

## Installing the package

```{r, eval=F}
## install and load devtools
install.packages("devtools")
library(devtools)

##install and load sExtinct from github
install_github("ExperimentalConservation/sExtinct")
library(sExtinct)

```

## Example analysis

```{r, echo=F, message=F}
library(sExtinct)
```

Example data are provided within the package:

```{r}
## load in the example data set
data("example.data")

##series of sightings 
print(example.data)
```

`Run.all` runs all of the included sightings based indicators simultaneously. Used must specify values for `alpha` (1-alpha confidence interval), `test.year` (typically set to the current year, i.e. test whether the species is currently predicted to be extant or extinct), whether to return the data, and whether the plot the results:

```{r}
run.all(example.data,
        alpha = 0.05,
        test.year = 2023,
        data.out = F,
        plot = T)

```


# References

[Clements, C., Collen, B., Blackburn, T., Petchey, O. Historic environmental change may affect our ability to infer extinction status. Conservation Biology, 28: 971–981.](https://conbio.onlinelibrary.wiley.com/doi/abs/10.1111/cobi.12329)

# Contact
Chis Clements - c.clements@bristol.ac.uk
